Containers and R---the Rockerverse and beyond

### 2020-01-09 manuscript.Rmd
add idea to conclusion, closes #2
add Rockerverse hex sticker

@@

- next steps
  - consolidation?
- support more complex set-ups with ready-to-use stacks, e.g. GPU, AI, ML
- Missing pieces?
- consolidation (e.g. via packages using `dockerfiler` and `stevedore`)
- Common themes
  - reproducibility
- will knowledge about containers continue to spread?
- what is needed for even more containers with R?
- the ability to move processing between services easily (e.g locally, one cloud providers VM, another cloud provider's Container-as-a-Service)
- running RStudio in a container is a common approach for reproducibility - _is there a need to run/control containers from a UI integrated in RStudio?_
- moving away from Docker focus towards containers?
  - is there a need for a wrapper for [podman's Varlink-API](https://github.com/containers/libpod/blob/master/API.md) (similar to stevedore) and buildah to manage containers and images respectively with alternative software stacks?
  - will there be control packages for Singularity support?
- ...

@@

>>>>

```{r hexSticker, message=TRUE, warning=FALSE, cache=TRUE, include=FALSE}
if(!file.exists("rockerverse.png")) {
  library("magick")
  library("here")

  # https://commons.wikimedia.org/wiki/File:NASA-HS201427a-HubbleUltraDeepField2014-20140603.jpg
  universe <- magick::image_read("https://upload.wikimedia.org/wikipedia/commons/6/69/NASA-HS201427a-HubbleUltraDeepField2014-20140603.jpg")
  magick::image_write(universe, here("universe.png"))
  # rockerLogoMask <-
  magick::image_read("https://avatars2.githubusercontent.com/u/9100160?s=200&v=4") %>%
    magick::image_scale("2100x2100") %>%
    magick::image_channel("Matte") %>%
    image_negate() %>%
    magick::image_border("black", "150x0") %>%
    magick::image_write(here("rocker_mask.png"))

  # does not work: image_composite(universe, rockerLogoMask, operator = "CopyOpacity", compose_args = "-composite -alpha Off")
  system2("convert", c(here("universe.png"), here("rocker_mask.png"),
    "-alpha Off",
    "-compose CopyOpacity",
    "-composite",
    here("whaleverse.png")))

  library("hexSticker")
  hexSticker::sticker(here("whaleverse.png"), package = "Rockerverse",
          p_size = 20, s_x = 0.97, s_y = 0.95, s_width = 1,
          h_fill = "#000000", h_color = "#ffffff",
          url = "github.com/nuest/rockerverse-paper",
          u_color = "#ffffff", u_size = 4,
          filename = here("rockerverse.png"))

  file.remove(c(here("universe.png"), here("whaleverse.png"), here("rocker_mask.png")))
}
```

![Rockerverse hex sticker](`r here::here("rockerverse.png")`)

@@

- what are next steps?
- support more complex set-ups with ready-to-use stacks, e.g. GPU, AI, ML
- Missing pieces?
- consolidation (e.g. via packages using `dockerfiler` and `stevedore`)
- Common themes
  - reproducibility
- will knowledge about containers continue to spread?
- what is needed for even more containers with R?
- the ability to move processing between services easily (e.g locally, one cloud providers VM, another cloud provider's Container-as-a-Service)
- running RStudio in a container is a common approach for reproducibility - _is there a need to run/control containers from a UI integrated in RStudio?_
- moving away from Docker focus towards containers?
  - is there a need for a wrapper for [podman's Varlink-API](https://github.com/containers/libpod/blob/master/API.md) (similar to stevedore) and buildah to manage containers and images respectively with alternative software stacks?
  - will there be control packages for Singularity support?
- sustain the overview and structure in this article in form of a CRAN Task View

@@

# Acknowledgements

DN is supported by the project Opening Reproducible Research ([o2r](https://www.uni-muenster.de/forschungaz/project/12343)) funded by the German Research Foundation (DFG) under project number [PE&nbsp;1632/17-1](https://gepris.dfg.de/gepris/projekt/415851837).
The funders had no role in data collection and analysis, decision to publish, or preparation of the manuscript.
